Using a Highly Dynamic Language for Development

### 2021-10-11 gic2021-highly-dynamic/paper.tex
minor

\section{Modularity Through Mixins}
The Common Lisp Object System (CLOS) has a couple of traits that remain rare in programming languages in use today, but make for excellent tools to support game development. Relevant to this section are serialised multiple inheritance and the standard generic function method combination.

@@

While the standard method combination and mixins offer a great deal of flexibility out of the gate, great care must still be taken when designing the overall class hierarchy and function protocol. Otherwise behaviours will not combine cleanly and lead to strange bugs, or might not combine properly at all.

@@

Restarts are on first look similar to conditions and handlers. When a restart is established, it sets an unwind point for a particularly named restart with a function to be called when that restart is triggered. Restarts themselves can be triggered via a function called \code{invoke-restart}, which can also pass along arguments for the restart function. The idea with restarts is to provide one or more ways for execution to continue or recover safely from an error. When the debugger is invoked from an unhandled error, it now has access to a list of active restarts on the stack, and can invoke one of them dynamically as the user sees fit.

@@

\begin{listing}[H]
\begin{minted}[fontsize=\scriptsize]{common-lisp}
(defmethod render ((scene scene))
  (for ((object in scene))
    (with-simple-restart (continue)
      (render object))))

@@

However, importantly, for the \code{main} \code{render}\\ method, this restart is automatically invoked, ensuring that any potential crashes at runtime of the game are gracefully ignored. This handler could also be defined elsewhere even higher up the stack, or only invoke the restart under special conditions.

>>>>

\section{Modularity Through Mixins}
The Common Lisp Object System (CLOS) has a couple of traits that remain rare in programming languages in use today, but make for excellent tools to support game development. Relevant to this section are serialised multiple inheritance and the standard generic function method combination. These features can be used to implement a system resembling ECS, though also offer a few advantages that make combination of behaviours more natural.

@@

While the standard method combination and mixins offer a great deal of flexibility out of the box, great care must still be taken when designing the overall class hierarchy and function protocol. Otherwise behaviours will not combine cleanly and lead to strange bugs, or might not combine properly at all.

@@

Restarts are on first look similar to conditions and handlers. When a restart is established through \code{restart-case}, it sets an unwind point for a particularly named restart with a function to be called when that restart is triggered. Restarts themselves can be triggered via a function called \code{invoke-restart}, which can also pass along arguments for the restart function. The idea with restarts is to provide one or more ways for execution to continue or recover safely from an error. When the debugger is invoked from an unhandled error, it now has access to a list of active restarts on the stack, and can invoke one of them dynamically as the user sees fit.

@@

\begin{listing}[H]
\begin{minted}[fontsize=\scriptsize]{common-lisp}
(defmethod render ((scene scene))
  (for ((object in scene))
    (restart-case
        (render object)
      (continue ())))

@@

However, importantly, for the \code{main} \code{render}\\ method, this restart is automatically invoked, ensuring that any potential crashes at runtime of the game are gracefully ignored. This handler could also be defined elsewhere even higher up the stack, or only invoke the restart under special conditions. This works as the \code{lambda} of the \code{handler-bind} is invoked at the point of error on top of the stack, where the \code{continue} restart is visible.
